---
title: "Lake Superior Scheduler Instructions"
author: "Derek H. Ogle"
date: '`r format(lubridate::now(),format="%d-%b-%Y")`'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Scheduler Introduction
Blah-blah-blah

# Scheduler Setup
## Scheduler Program Files
Blah-blah-blah

## R Packages
The scheduler (and creel analysis programs require a number of packages that may not be already installed in your R. These packages can be installed by running the following code in RStudio. You should only have to run this code once on each computer.

```{r eval=FALSE,prompt=FALSE}
rqrd <- c("captioner","devtools","dplyr","FSA","ggplot2","haven","here",
          "huxtable","kableExtra","knitr","lubridate","tidyr","magrittr",
          "RColorBrewer","readxl","rlang","rmarkdown","sugrrants","tibble")
utils::install.packages(rqrd)
devtools::install_github("thomasp85/patchwork")
```

## RStudio Project
The scheduler program depends on you making an "RStudio Project" for the specific folder at the root of the file structure created [above](#analysis-program-files) (i.e., the folder in which you created the "LSuperior" folder). RStudio projects can be easily created by opening RStudio, selecting the "File" menu item, selecting "New Project ...", selecting "Existing Directory", then browsing to the root folder that contains the creel analysis programs, and selecting "Create Project". RStudio will then move to the project and you should see the name of the root folder just below the toolbar in the upper-right corner of the RStudio window.

[SCREEN CAPTURE HERE]

## Scheduler Information File
Blah-blah-blah


# Scheduler Assumptions
The Lake Superior open-water creel clerk scheduler and bus route constructor is built upon the following premises and assumptions.

1. There are four creel clerks:
    * **Clerk A**: Ashland-Washburn (100%; full-time)
    * **Clerk B**: Bayfield-Red Cliff-Little Sand Bay (50%) and Cornucopia-Port Wing (50%) (full-time)
    * **Clerk C**: Saxon (100%; part-time)
    * **Clerk D**: Superior (100%; full time)
1. Details for each clerk and route are in the "Schedule_Info_LS.xlsx" file.
1. Each clerk will conduct a survey on every weekend day and holiday during the "open-water" months. Holidays during the "open-water" months are Memorial Day, Independence Day (i.e., Fourth of July), and Labor Day.
1. Each clerk will have two consecutive weekdays off each week during the "open-water" months. The same two weekdays will not be off for consecutive weeks. Further, if the clerk had Thursday-Friday off one week then they will not have Monday-Tuesday off the following week (*this limits the number of "only two consecutive days worked" occurrences*) and if the clerk had Monday-Tuesday off one-week then they will not have Thursday-Friday off the following week (*this limits the number of "eight and nine consecutive days worked" occurrences*). Furthermore, the clerk is not allowed to have one weekday off between a holiday (that falls on a weekday) and the weekend.
    * Note that this tends to produce a disproportionate amount of Tuesdays, Wednesdays, and Thursdays as "off days."
1. Shifts (am/pm) are created so that there are roughly (within one or two) even numbers of am and pm shifts in each month for each route. However, the shifts are randomly placed in the days such that there may be a (randomly generated) run of the same shift for several days in a row.
1. For the clerk responsible for two routes (i.e., B).
    * The route for a weekend/holiday is chosen by randomly selecting the route for the first weekend day or holiday in a two-week period, using the other route for the second weekend day or holiday, then switching the order for the next two weekend days or holidays, and then randomly selecting a route if there is a fifth weekend day or holiday in the two-week period. Thus, approximately the same number of routes are run for the weekends and holidays, but the routes will not be run on the same weekend days for more than two consecutive weekends (except for some rare occasions in weeks with holidays).
    * The route for a weekday is chosen by randomly selecting the route for the first weekday in a two-week period and then alternating routes after that for each weekday in the rest of the two-week period. Thus, the same number of routes is run for the weekdays of each two-week period, but the routes will not be run on the same weekdays for more than two consecutive weeks.
